---
title: "TWINSPAN in classification comparison - R Notebook"
output: html_notebook
date: "Last Rendered on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    fig_width: 7
    fig_height: 7
always_allow_html: true 
---






## Libraries
Load necessary libraries
```{r}
# # NB if you need to install twinspan you need to RUN R/RSTUDIO AS ADMINSTRATOR
# library(devtools)
# devtools::install_github("jarioksa/twinspan")

#option 2 if that fails
#install.packages("twinspan", repos="https://jarioksa.github.io/drat/")

###

library(twinspan)
library(tidyverse)
#library(rgdal)
library(sf)
#library(rgeos)
```

## Load input samples by species data (made by Jonatan in script 01_Video_DataPrep.Rmd here: https://github.com/JonBrox/MAREANO_Video/blob/0dbff93e13319cfd86f2ca4af9cb49285cddc03c/01_Video_DataPrep.Rmd)
```{r}
#load otu file
load(file="C:/Users/a21448/Havforskningsinstituttet/MAREANO Biotope modelling - General/VideoData/Output/otu_Video_All.rda")

#load env file
load(file="C:/Users/a21448/Havforskningsinstituttet/MAREANO Biotope modelling - General/VideoData/Output/env_Video_All.rda")

#TEMPORARY
#steal the SampID column from env and add it to otu

otuwide<-cbind(env$SampID, otu) %>% select(order(colnames(.))) %>% rename("SampID"="env$SampID")

otuwide<-otuwide%>%select(SampID, everything())

#load sample info table
sample_info<-read.csv("C:/Users/a21448/Havforskningsinstituttet/MAREANO Biotope modelling - General/VideoData/Input/sample_info.csv")

```


## REDUCE TO OUR AREA

```{r}
#load Rnumber table (dbf > csv from Geno's shapefile)
Rsel<-read.csv("C:/Users/a21448/Havforskningsinstituttet/MAREANO Biotope modelling - General/VideoData/Input/R_stations_comparisonStudy.csv")

# use Rsel$smpl_ to select VLs from sample_Info$VL and use that to select SampID from trimmedwide
subSampleInfo<-sample_info[sample_info$VL %in% Rsel$smpl_,]
trimmedwide<-otuwide[otuwide$SampID %in% subSampleInfo$SampID,]


```



## EXPLORATIONS

### Check species included
```{r}
colnames(trimmedwide)
```

### check number of species per sample
```{r}
sppPerSamp<-trimmedwide %>% mutate(noSpp=rowSums(select_if(., is.numeric)!=0)) %>% select(c("SampID", "noSpp"))%>% arrange(noSpp)
#MAX NO OF SPP PER SAMP
message("Maximum number of species per sample")
max(sppPerSamp$noSpp)
message("Minimum number of species per sample")
min(sppPerSamp$noSpp)

plot(sppPerSamp$noSpp)
```
## Inspect samples with few spp.
```{r}
fewSppPerSamp<-subset(sppPerSamp,noSpp<2)
fewSppSamp<-trimmedwide[trimmedwide$SampID %in% fewSppPerSamp$SampID,]
fewSppSamp<-fewSppSamp %>%
 select(where(~ any(. != 0)))
# table of one spp Samples - which species are present - any consistency?
fewSppSamp
#- nope! 
```





### check number of samples per species
```{r}
sampPerSp<-as.data.frame(colSums(trimmedwide[2:175]!=0)) %>% rename("noSamp"="colSums(trimmedwide[2:175] != 0)") %>% arrange(noSamp)
message("Maximum number of samples per species")
max(sampPerSp$noSamp)
message("Minimum number of samples per species")
min(sampPerSp$noSamp)
```


### check the mean/max density per sample
```{r}
 densMaxMean<-trimmedwide %>% rowwise() %>% mutate (
   max=max(c_across(Acesta_excavata:Tubularia)),
   mean=mean(c_across(Acesta_excavata:Tubularia))) %>% select(SampID, max, mean)

plot(densMaxMean$max)
plot(densMaxMean$mean)
```


## Check where samples are
```{r}
trWmeta<-left_join(trimmedwide, sample_info)

plot(trWmeta$x_coord, trWmeta$y_coord)
```

# TWINSPAN TIME


## Threshold the number of spp/samp samp/spp

```{r}
# minNoSppPerSamp<-4
# minNoSampPerSpp<-4


```


## Perform the twinspan analysis

### Fit the twinspan model
Turn on or off the filters as you wish
```{r, warning=FALSE}

samples <- trimmedwide 

coredata <- trimmedwide
row.names(coredata) <- samples$SampID
coredata <- coredata %>% select(-SampID)


#cutlevels = c(0,100,200,400,1000)
cutlevelsPaal = c(0,1,2,5,10,20)

tw_1 <- coredata %>%
  select(which(!is.na(colSums(.)))) %>%
  #mutate_all(sqrt)%>%
  twinspan(cutlevels = cutlevelsPaal, levmax = 15, groupmin = 4) #change groupmin to 4? (see minNoSampPerSpp)

#summary(tw_1)
eval <- misclassified(tw_1)
eval <- with(eval,  data.frame(SampID = samples$SampID[index], class = class, predicted = predicted))
eval
```
## Save important variables
```{r}
save(tw_1,eval, file="twinspResults_compArea492obs.rda")
```





## check excluded samples/stations
outputting for more detailed explorations
```{r}
# # excluded<-refer %>% filter(sample_no%in%unique(sample_info$VL)) %>%
# #   filter(!(sample_no%in%good))
# # 
# # write.csv(excluded, file = file.path(outPath, ("excluded_megaAnalysis.csv")))
# #   
# 
# excl.samp<-sample_info %>%
#   anti_join(tw_1.df, by=c("SampID"="rowname"))%>%
#   left_join(refer, by=c("VL"="sample_no"))
# 
# write.csv(excl.samp, file = file.path(outPath, ("excluded_samples.csv")))
